Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds

### 2015-12-23 paper.tex
final all-in version

\title
[Testing Quasar Unification: Radiative Transfer in Clumpy Winds]{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Testing Quasar Unification: Radiative Transfer in Clumpy Winds
%Unifying quasars with clumpy wind models
}

@@

\begin{abstract} 
Various unification schemes interpret the complex phenomenology of quasars and luminous active galactic nuclei (AGN) in terms of a simple picture involving a central black hole, an accretion disc and an associated outflow. Here, we continue our tests of this paradigm by comparing quasar spectra to synthetic spectra of biconical disc wind models, produced with our state-of-the-art Monte Carlo radiative transfer code. Previously, we have shown that we could produce synthetic spectra resembling those of observed broad absorption line (BAL) quasars, but only if the X-ray luminosity was limited to about 
10$^{43}$~erg~s$^{-1}$. Here, we show that clumping the outflow can maintain the ionization state necessary for BAL features in the rest-frame UV at more realistic X-ray luminosities. Our fiducial model shows good agreement with X-ray properties of AGN and the dense, X-ray heated wind produces strong line emission in, e.g., \la\ and \civline at low inclinations. At high inclinations, the synthetic spectra possess prominent LoBAL features. Despite these successes, we cannot reproduce all the emission lines seen in quasar spectra with the correct equivalent-width ratios, and and we find an angular dependence of emission-line equivalent width despite the remarkable similarities in the observed emission line properties of BAL and non-BAL quasars. Overall, our work suggests that a biconical wind can reproduce much of the qualitative behaviour expected from a unified model, but we cannot yet provide quantitative matches with quasar properties at all viewing angles. Whether disc winds can successfully unify quasars is therefore still an open question.
\end{abstract}

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}),
which are reprocessed by `macro-atoms' with associated transition probabilities.
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15. In this scheme, the energy packets interact with either two-level 
`simple atoms' or full macro-atoms. 
This allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple atoms (see M15). In this study,
only H and He are treated as macro-atoms, because the process is computationally
intensive, and we expect recombination to be particularly important
in determining their level populations and resultant line emission.
We are also especially interested in the contribution to 
AGN spectra of \LA.  H13 treated all atoms in a two-level approximation.

@@

\begin{figure} 
\centering
\includegraphics[width=0.45\textwidth]{figures/fig2_cartoon.png}
\caption
{
A cartoon showing the geometry and some key parameters of
our biconical wind model.
}
\label{fig:cartoon}
\end{figure}

@@

The X-ray source is treated as an isotropic sphere at the ISCO,
which emits $r$-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K_X \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
Photons, or $r$-packets, produced by the accretion disc and central X-ray source
are reprocessed by the wind. This reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet constraint (see Lucy 2002, M15).  
% In addition to the disc and X-ray source, 
% the wind is able to reprocess radiation. However, new 
% photon packets are not produced in the wind (as in LK02). 
% Instead, this reprocessing is dealt with by enforcing strict
% radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
% via an indivisible energy packet
% constraint (see Lucy 2002, M15).  
% {\bf I confess, I do not understand what is being said here.  I would like to understand it better, and I would like it explained better.  How do we avoid, generating for example recombination photons from non-macro atoms.}

@@

\section{Results and Discussion From A Fiducial Model}
Here we describe the results from a fiducial model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in Table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN. 
% We also examine trends with inclination in the synthetic spectra, 
% both in terms of the range of ionization states of the absorption lines and equivalent widths of the emission lines.

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Fiducial Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{m}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{erg~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{m}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
%%$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f_V$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters 
used in the fiducial model, as defined in the text and figure 1.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk.}
\label{wind_param}
\end{table}

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/link8.png}
\caption
{
Contour plots showing the logarithm of some important 
physical properties of the outflow. The spatial scales are
logarithmic and the $x$ and $z$ scales are not the same.
Symbols are defined in the text.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
The line luminosities, $L$, represent the luminosity of photons
escaping the Sobolev region for each line. These photons do not
necessarily escape to infinity.
}
\label{fig:wind}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles for the fiducial model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
SDSS J162901.63+453406.0, respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled to $1350$\AA\
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure} %fullpage

\begin{figure} 
\centering
\includegraphics[width=0.5\textwidth]{figures/windnew3.png}
\caption
{
A cartoon describing the broad classes of sightline 
in the fiducial model, illustrating how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/sed.png}
\caption
{
Synthetic spectra at two viewing angles, 
this time in frequency space and including the optical band,
compared to the non-BAL SDSS quasar composite. The spectra are normalised to the flux at 
$2000$\AA, then an offset of 2 is applied per spectrum for clarity -- the dotted lines show the zero point of $F_\nu / F_{2000\AA}$ in each case.
}
\label{fig:sed}
\end{figure} %fullpage

@@

\begin{table}
\begin{tabular}{p{2cm}p{2cm}p{3cm}}
\hline Property & Synthetic, $20^\circ$ & Observed  (S11) \\ 
\hline \hline
$\log L$[C~\textsc{iv}]  & $44.60$ & $44.42 \pm 0.32$  \\
$\log L$[Mg~\textsc{ii}] & $43.92$ & $43.54 \pm 0.28$  \\
$\log (\nu L_{\nu})_{1350}$  & $46.42$ & $46.01 \pm 0.30$ \\
$\log (\nu L_{\nu})_{3000}$  & $46.18$ & $45.79 \pm 0.30$ \\
% $\log L$[H~$\alpha$] & $43.46$ & $42.86$ & ?? \\
% $EW$[C~\textsc{iv}] (erg~s$^{-1}$) & $3.96 \times 10^{44}$ & $2.62 \times 10^{44}$  \\
% $EW$[Mg~\textsc{ii}] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% $EW$[H~$\alpha$] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% $\log L_{1350}$ (erg~s$^{-1}$~Hz$^{-1}$) & $46.42$ & $45.56$ & $46.01 \pm 0.30$ \\
% $\log L_{3000}$ (erg~s$^{-1}$~Hz$^{-1}$) & $46.18$ & $45.36$ & $45.79 \pm 0.30$ \\
% $L_{bol})$ (erg~s$^{-1}$) & 2.62
% % $L_{1300})$(C\textsc{iv}) & 
% \\
% \multicolumn{3}{|l|}{At $73^\circ$ (`BALQSO-like')} \\  
% $\log L$[C~\textsc{iv}] (erg~s$^{-1}$) & $44.14$ & ??  \\
% $\log L$[Mg~\textsc{ii}] (erg~s$^{-1}$) & $42.82$ & $44.41$  \\
% $\log L$[H~$\alpha$] (erg~s$^{-1}$) & $42.86$ & ?? \\
% % $EW$[C~\textsc{iv}] (erg~s$^{-1}$) & $3.96 \times 10^{44}$ & $2.62 \times 10^{44}$  \\
% % $EW$[Mg~\textsc{ii}] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% % $EW$[H~$\alpha$] (erg~s$^{-1}$) & $3.50 \times 10^{43}$ & $8.32 \times 10^{43}$ \\
% $\log L_{1350}$ (erg~s$^{-1}$~Hz$^{-1}$) & $45.56$ & $46.00$ \\
% $\log L_{3000}$ (erg~s$^{-1}$~Hz$^{-1}$) & $45.36$ & $45.79$ \\
\hline
\end{tabular}
\caption{
Some derived spectral properties of the fiducial model, at $20^\circ$,
compared to observations. The observed values are taken from the Shen et al. (2011)
SDSS DR7 Quasar catalog, and correspond to mean values with standard deviations in log space
from a subsample with $8.5>\log(M_{BH})<9.5$ and $1.5<\log (L_{bol}/L_{Edd}) < 0$,
where the BH mass is a \civ\ virial estimate. 
Units are logarithms of values in erg~s$^{-1}$.
%unless stated.otherwise.
}
\label{line_lums}
\end{table}

@@

In Fig.~\ref{fig:sed}, we show an $F_{\nu}$ spectrum with broader waveband coverage
that includes the optical, showing that our synthetic spectra 
also exhibit \ha\ and \hb\ emission. 
In this panel, we include a low inclination and 
also a very high inclination 
spectrum, which looks underneath the wind cone. This model shows 
strong line emission with very similar widths and line ratios to the quasar composites, and
the Balmer lines are double peaked, due to velocity projection effects.  
Such double-peaked lines are seen in so-called `disc emitter' systems 
\citep[e.g.][]{eracleous1994} but not the majority of AGN.     
The line equivalent widths (EWs) increase at high inclination
due to a weakened continuum from wind attenuation, 
disc foreshortening and limb darkening. This effect also 
leads to a redder continuum slope, as seen in quasars, which is
due to Balmer continuum and Balmer and Fe~\textsc{ii} line emission.
This extreme $89^\circ$ viewing angle cannot represent a typical quasar within a unified model,
but does show that a model such as this can naturally reproduce quasar emission lines
if the emissivity of the wind is increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.
% This is because we do not fully solve the C level populations,
% and thus the level populations do not accurately reflect those
% as a full macro-atom with a 
% full collisional rates between forbidden
% or semi-forbidden transitions, as would be required.
% The critical density of the C~\textsc{iii}]~1909\AA\, line 
% is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
% outer portion of our wind. 
% We therefore expect a model with these 
% parameters to produce a C~\textsc{iii}]~1909\AA\ line
% with a proper treament. We note, however, the the presence of
% this line in the spectrum of quasars does {\em not} prohibit denser material having a significant contribution to the spectrum.

% The model produces strong emission lines in \civ, \nv\ and \la,
% as well as a very weak \mg\ line.

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=0.9\textwidth]{figures/lx_a05_pre.png}
\caption
{
X-ray ($2$~keV) luminosity of the our clumped model (red squares) 
and the H13 model (blue squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme (see figure 4).
Also plotted are masurements from 
the COMBI-7 AGN and the BQS samples (Steffen et al. 2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted line shows the best fit relation for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/c4_angles.png}
\caption
{
\civ , \mg , \al\ and Fe~\textsc{ii} line profiles for viewing angles
from $72-79^\circ$. The profiles are plotted relative to the local
continuum with an offset applied for clarity. Lower ionization
profiles appear at a subset of high inclinations, compared
to the ubiquitous \civ\ profile.
}
\label{fig:lobal}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/scatter_c4_grid.png}
\caption
{
The EW of the \civ~$1550$\AA\ line at $20^\circ$ plotted against a) the 
$BI$ of \civ~$1550$\AA\ at $75^\circ$, b) the EW of the \mg~$2800$\AA\ line 
at $20^\circ$ and c) the EW of \la\ at $20^\circ$. The circles correspond 
to the simulation grid for two different values of $f_V$, and the fiducial 
model is marked with an orange star. 
We also show a higher X-ray luminosity model and a higher mass loss rate
with a red star.
}
\label{fig:grid}
\end{figure} %fullpage

@@

\bibliography{mybib.bib,stellar.bib,h14.bib,h13.bib,hamann.bib,krolik.bib}
\clearpage
\clearpage
%\appendix
% \appendix
% \section{Notes}

>>>>

\title
[Radiative Transfer in Clumpy Winds]{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Testing Quasar Unification: Radiative Transfer in Clumpy Winds
%Unifying quasars with clumpy wind models
}

@@

\begin{abstract} 
Various unification schemes interpret the complex phenomenology of quasars and luminous active galactic nuclei (AGN) in terms of a simple picture involving a central black hole, an accretion disc and an associated outflow. Here, we continue our tests of this paradigm by comparing quasar spectra to synthetic spectra of biconical disc wind models, produced with our state-of-the-art Monte Carlo radiative transfer code. Previously, we have shown that we could produce synthetic spectra resembling those of observed broad absorption line (BAL) quasars, but only if the X-ray luminosity was limited to
10$^{43}$~erg~s$^{-1}$. Here, we show that clumping the outflow can maintain the ionization state necessary for BAL features in the rest-frame UV at more realistic X-ray luminosities. Our fiducial model shows good agreement with X-ray properties of AGN and the dense, X-ray heated wind produces strong line emission in, e.g., \la\ and \civline at low inclinations. At high inclinations, the synthetic spectra possess prominent LoBAL features. Despite these successes, we cannot reproduce all the emission lines seen in quasar spectra with the correct equivalent-width ratios, and and we find an angular dependence of emission-line equivalent width despite the remarkable similarities in the observed emission line properties of BAL and non-BAL quasars. Overall, our work suggests that a biconical wind can reproduce much of the qualitative behaviour expected from a unified model, but we cannot yet provide quantitative matches with quasar properties at all viewing angles. Whether disc winds can successfully unify quasars is therefore still an open question.
\end{abstract}

@@

Our approach to line transfer is based upon the macro-atom implementation developed by 
\cite{lucy2002, lucy2003}, in which the energy flows through the system are described in 
terms of indivisible energy quanta of radiant or kinetic energy 
(`$r-$packets' and `$k-$packets' respectively; see also section~\ref{sec:photon_sources}),
which are reprocessed by `macro-atoms' with associated transition probabilities.
We use the Sobolev approximation \citep[e.g.][]{sobolev1957,sobolev1960,rybickihummer1978}
to compute the location of line interactions.
In our case, for reasons of computational efficiency, we adopt the  hybrid macro-atom scheme 
described by M15. In this scheme, the energy packets interact with either two-level 
`simple atoms' or full macro-atoms. 
This allows one to treat non-LTE line transfer in radiative equilibrium
without approximation for elements that are identified as 
full macro-atoms, while maintaining the fast `two-level' 
treatment of resonance lines when elements are identified 
as simple atoms (see M15). In this study,
only H and He are treated as macro-atoms, because the process is computationally
intensive, and we expect recombination to be particularly important
in determining their level populations and resultant line emission.
We are also especially interested in the contribution to 
AGN spectra of \LA.  H13 treated all atoms in a two-level approximation.

@@

\begin{figure} 
\centering
\includegraphics[width=0.45\textwidth]{figures/fig1.eps}
\caption
{
A cartoon showing the geometry and some key parameters of
our biconical wind model.
}
\label{fig:cartoon}
\end{figure}

@@

The X-ray source is treated as an isotropic sphere at the ISCO,
which emits $r$-packets according to a power law in flux with index $\alpha_X$, of the form
\begin{equation}
F_X (\nu) = K_X \nu^{\alpha_X}.
\end{equation}
The normalisation, $K_X$ of this power law is such that it 
produces the specified 2-10~keV luminosity, $L_X$.
Photons, or $r$-packets, produced by the accretion disc and central X-ray source
are reprocessed by the wind. This reprocessing is dealt with by enforcing strict
radiative equilibrium ({\em modulo} adiabatic cooling; see section~2.3)
via an indivisible energy packet constraint (see Lucy 2002, M15).

@@

\section{Results and Discussion From A Fiducial Model}
Here we describe the results from a fiducial model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in Table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN.

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Fiducial Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{m}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{erg~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{m}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
%%$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.5^*$ \\
$f_V$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\end{tabular}
\caption{Wind geometry parameters 
used in the fiducial model, as defined in the text and figure 1.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk.}
\label{wind_param}
\end{table}

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/fig2.eps}
\caption
{
Contour plots showing the logarithm of some important 
physical properties of the outflow. The spatial scales are
logarithmic and the $x$ and $z$ scales are not the same.
Symbols are defined in the text.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
The line luminosities, $L$, represent the luminosity of photons
escaping the Sobolev region for each line. These photons do not
necessarily escape to infinity.
}
\label{fig:wind}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/fig3.eps}
\caption
{
Synthetic spectra at four viewing angles for the fiducial model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
SDSS J162901.63+453406.0, respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled to $1350$\AA\
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure} %fullpage

\begin{figure} 
\centering
\includegraphics[width=0.5\textwidth]{figures/fig4.eps}
\caption
{
A cartoon describing the broad classes of sightline 
in the fiducial model, illustrating how geometric effects lead to 
the different emergent spectra. The colour gradient is approximate,
but indicates the stratified ionization structure, 
from highly ionized (yellow) to low ionization (purple) material.
%[FIGURE NEEDS IMPROVING]
}
\label{fig:sightline}
\end{figure}

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/fig5.eps}
\caption
{
Synthetic spectra at two viewing angles, 
this time in frequency space and including the optical band,
compared to the non-BAL SDSS quasar composite. The spectra are normalised to the flux at 
$2000$\AA, then an offset of 2 is applied per spectrum for clarity -- the dotted lines show the zero point of $F_\nu / F_{2000\AA}$ in each case.
}
\label{fig:sed}
\end{figure} %fullpage

@@

\begin{table}
\begin{tabular}{p{2cm}p{2cm}p{3cm}}
\hline Property & Synthetic, $20^\circ$ & Observed  (S11) \\ 
\hline \hline
$\log L$[C~\textsc{iv}]  & $44.60$ & $44.42 \pm 0.32$  \\
$\log L$[Mg~\textsc{ii}] & $43.92$ & $43.54 \pm 0.28$  \\
$\log (\nu L_{\nu})_{1350}$  & $46.42$ & $46.01 \pm 0.30$ \\
$\log (\nu L_{\nu})_{3000}$  & $46.18$ & $45.79 \pm 0.30$ \\
\hline
\end{tabular}
\caption{
Some derived spectral properties of the fiducial model, at $20^\circ$,
compared to observations. The observed values are taken from the Shen et al. (2011)
SDSS DR7 Quasar catalog, and correspond to mean values with standard deviations in log space
from a subsample with $8.5>\log(M_{BH})<9.5$ and $1.5<\log (L_{bol}/L_{Edd}) < 0$,
where the BH mass is a \civ\ virial estimate. 
Units are logarithms of values in erg~s$^{-1}$.
%unless stated.otherwise.
}
\label{line_lums}
\end{table}

@@

In Fig.~\ref{fig:sed}, we show an $F_{\nu}$ spectrum with broader waveband coverage
that includes the optical, showing that our synthetic spectra 
also exhibit \ha\ and \hb\ emission. 
In this panel, we include a low inclination and 
also a very high inclination 
spectrum, which looks underneath the wind cone. This model shows 
strong line emission with very similar widths and line ratios to the quasar composites, and
the Balmer lines are double peaked, due to velocity projection effects.  
Such double-peaked lines are seen in so-called `disc emitter' systems 
\citep[e.g.][]{eracleous1994} but not the majority of AGN.     
The line equivalent widths (EWs) increase at high inclination
due to a weakened continuum from wind attenuation, 
disc foreshortening and limb darkening. This effect also 
leads to a redder continuum slope, as seen in quasars, which is
due to Balmer continuum and Balmer and Fe~\textsc{ii} line emission.
This extreme $89^\circ$ viewing angle cannot represent a typical quasar within a unified model,
but does show that a model such as this can naturally reproduce quasar emission lines
if the emissivity of the wind is increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=0.9\textwidth]{figures/fig6.eps}
\caption
{
X-ray ($2$~keV) luminosity of the our clumped model (red squares) 
and the H13 model (blue squares), plotted against monochromatic luminosity 
at 2500\AA. The points are labeled according to inclination; angles
$>70^\circ$ correspond to BALs in our scheme (see figure 4).
Also plotted are masurements from 
the COMBI-7 AGN and the BQS samples (Steffen et al. 2006) and the Saez et al. (2012) 
sample of BALQSOs. The dotted line shows the best fit relation for non-BALQSOs 
from Steffen et al. (2006).
}
\label{fig:xray}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/fig7.eps}
\caption
{
\civ , \mg , \al\ and Fe~\textsc{ii} line profiles for viewing angles
from $72-79^\circ$. The profiles are plotted relative to the local
continuum with an offset applied for clarity. Lower ionization
profiles appear at a subset of high inclinations, compared
to the ubiquitous \civ\ profile.
}
\label{fig:lobal}
\end{figure} %fullpage

@@

\begin{figure} %fullpage
\centering
\includegraphics[width=1.0\textwidth]{figures/fig8.eps}
\caption
{
The EW of the \civ~$1550$\AA\ line at $20^\circ$ plotted against a) the 
$BI$ of \civ~$1550$\AA\ at $75^\circ$, b) the EW of the \mg~$2800$\AA\ line 
at $20^\circ$ and c) the EW of \la\ at $20^\circ$. The circles correspond 
to the simulation grid for two different values of $f_V$, and the fiducial 
model is marked with an orange star. 
We also show a higher X-ray luminosity model and a higher mass loss rate
with a red star.
}
\label{fig:grid}
\end{figure} %fullpage

@@

\bibliography{mybib.bib,stellar.bib,h14.bib,h13.bib,hamann.bib,krolik.bib}
\clearpage
\clearpage
